---
title: "Mendelian Histogram"
output: html_document
---

```{r, echo = F}
mendel <- read.table("mendel.result.c22.txt", header = T)


```

* Total number of rows in mendelian error data: `r nrow(mendel)`
* Total number of unique positions: `r length(unique(mendel$POS))`
* Total number of unique families: `r length(unique(mendel$FAMILY))`. Note that the total number of families in the original filtered data is 332 (996/3). 

```{r, echo = F}
a <- as.data.frame(sort(table(mendel$FAMILY)))
hist(mendel$POS, breaks = 40, xlab = "Positions", main = "Fig 1. Histogram of positions")
hist(a$Freq, main = "Fig 2. Histogram of positions by family", xlab = "# of positions", ylab = "# of family", breaks = 100)

```

## family outliners

```{r, echo = F}
head(sort(table(mendel$FAMILY), decreasing = T), 5)
```

## Try removing family outliners to check for positions
```{r, echo =F}
b <- head(sort(table(mendel$FAMILY), decreasing = T), 5)
mendel.family <- mendel[!mendel$FAMILY %in% names(b), ]

hist(mendel$POS, breaks = 40, xlab = "Positions", main = "Fig 1. Histogram of positions")
hist(mendel.family$POS, breaks = 40, xlab = "Positions", main = "Fig 3. Histogram of positions (family outliners removed)")

```

##Try removing position outliners to check for families

```{r, echo = F}
##region: 129340000 - 129360000, 129720000-129780000, 130160000-130180000
mendel.position <- mendel[! (mendel$POS >= 129340000  & mendel$POS <= 129360000), ]
mendel.position <- mendel.position[! (mendel.position$POS >= 129720000  & mendel.position$POS <= 129780000), ]
mendel.position <- mendel.position[! (mendel.position$POS >= 130160000  & mendel.position$POS <= 130180000), ]
c <- as.data.frame(sort(table(mendel.position$FAMILY)))

hist(a$Freq, main = "Fig 2. Histogram of positions by family", xlab = "# of positions", ylab = "# of family", breaks = 100)
hist(c$Freq, main = "Fig 4. Histogram of positions by family (position outliner removed)", xlab = "# of positions", ylab = "# of family", breaks = 100)
```

##check for extreme family by position
```{r}
target.position <- mendel[mendel$POS >= 129720000  & mendel$POS <= 129780000, ]
b <- as.data.frame(sort(table(target.position$FAMILY)))
hist(b$Freq, breaks = 30, main = "Fig 5. Histogram of family count for targeted positions", xlab = "# of positions", ylab = "# of family")

tail(sort(table(target.position$FAMILY)))



```

